Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-01-24 text/src/article.tex
Revise first section of results

% }}}
% {{{ Background
\section*{Background}
 Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
% }}}
% {{{ Implementation
%%%%%%%%%%%%%%%%%%%%%%
\section*{Implementation}
  Text for this section \ldots
% }}}
\section*{Results and Discussion} %{{{1
\subsection*{Scaffolder Description} %{{{2

  * Join contigs together
  * Add inserts
  * Add unknown gap regions
  * Specified in a YAML file format

\subsection*{Scaffold file format} %{{{2

@@

@@

  * Install through rubygems
  * Make sure the bin is in the class path

>>>>

% }}}
% {{{ Background
\section*{Background}
 Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
% }}}
% {{{ Implementation
%%%%%%%%%%%%%%%%%%%%%%
\section*{Implementation}
  Text for this section \ldots
% }}}
\section*{Results and Discussion} %{{{1

\subsection*{Scaffolder Description} %{{{2

Individual reads produced from nucleotide sequencing can be assembled into
contiguous regions (``contigs'') of many overlapping individual reads. These
reads may however not represent a complete genome sequence where all contigs do
not result in a single complete sequence. The software outlined here
``Scaffolder'' allows these contigs to be further joined together into
a continuous single sequence. This selection, order and direction of contigs
into a final draft genome is often described as the `scaffold'. Scaffolder
allows the manual or computational specification of a genome scaffold by
specifying which contig sequences are used in simple to edit configuration
files.

The scaffold file specifies how contigs should be joined together in the draft
genome sequence. The raw nucleotide sequence for each contig is stored in
a fasta file, as might be produced from assembly of sequencing reads using
programs such as newbler\cite{newbler} or abyss\cite{abyss}. Scaffolder then
the contig to be selected by specifying the name of the contig in the scaffold
file. The scaffold file also allows a sub-sequence or the reverse complement of
a contig to used. This therefore preserves the original sequence data but also
allowing sequences to manipulated in the final genome build.

The majority of a genome can be estimated from scaffolding assembled contigs
into the most parsimonious build. There may however still be internal gaps in
the contigs from mate pair assembly based on estimated mate pair insert size.
Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
and filled - a common process in finishing a genome sequence. Inserts may be
treated the same way as a contig where the raw sequence, as may be produced by
PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
in the contig.

Scaffolder allows the specification of unknown regions in the scaffold. These
are areas of the genome where the nucleotide sequence is unknown but the
approximate size region may be estimated from mate pair insert size or from
mapping contigs to reference genome and estimating the size of the gap. These
unknown regions allow a continuous draft sequence to be scaffolded but whilst
still highlighting regions waiting to be resolved.

\subsection*{Defining a scaffold in a text file} %{{{2

The genome scaffold is defined YAML format. YAML is an easy to edit data format
that is both machine readable and easy to edit manually.

INSERT FIGURE WITH EXAMPLE LAYOUT OF SCAFFOLD

@@

Scaffolder provides a ruby API that allows the scaffold and designated
sequences to be manipulated through a ruby programme. This allows scaffolder to
be integrated into exiting genomics works flows such as using Rake. Scaffolder
all provides a command line binary for the manipulation of the scaffold. This
binary behaves as a standard Unix command line interface returning appropriate
exit statuses and providing manual pages for each of the commands. These two
methods are outlined in the following two sections.

@@

  * Install through RubyGems
  * Make sure the bin is in the class path
